EldersAIR: data merging for analysis dataset

### 2021-05-07 manuscript_tables_figures.Rmd
fibit data cleaning new interaction analysis

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")))
```

@@

# summary data for manuscript tables - averaged over multiple Winter 1 visits
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  #filter(area == "NPT") %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(mean_var) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  #filter(area == "NPT") %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(treatment, mean_var) %>% 
  group_by(treatment) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")

>>>>

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

@@

# summary data for manuscript tables - averaged over multiple Winter 1 visits
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  #filter(area == "NPT") %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(mean_var) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  #filter(area == "NPT") %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(treatment, mean_var) %>% 
  group_by(treatment) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")
num_funct_total("mp_mean_winter")
num_funct_treatment("mp_mean_winter")

### 2021-05-07 merge_data_files.Rmd
fibit data cleaning new interaction analysis

---
title: 'EldersAIR: data merging for analysis dataset'
author: "Ethan Walker"
date: "Started 11 March 2021, Updated 15 March 2021"
output:
  powerpoint_presentation: default
---

@@

elders_at_home_log_clean <- read_rds(paste0(file_path, "Output/elders_at_home_log_clean.rds"))

@@

elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds")) 
```

Merge PM and at-home data and summarize by sampling visit
```{r}
elders_at_home_log <- elders_at_home_log_clean %>% 
  select(-home_winter_id, -home, -treatment)

@@

Merge PM, health, stove data 
One observation per sampling visit (up to 4 obs per adult/home)
```{r}
health_pm_visit <- elders_medvisit_clean %>% 
  full_join(elders_pm_at_home_visit,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(demographics,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_extra_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_hx_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(elders_stove_grades_clean,
            by = c("area", "home", "home_id_num", "treatment")) %>% 
  left_join(elders_stove_usage_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_kw_report_clean,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_filter_type,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_temp_rh_clean,
            by = c("area", "home_winter_id", "sampling_visit")) %>% 
  left_join(elders_home_activity_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_home_char_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(elders_moisture_clean,
            by = c("area", "home_winter_id", "sampling_visit"))

# summary(health_pm_visit)
```

Save final dataset
```{r}
write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))

write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")

# save variable names to use in data dictionary
save_var_names <- data.frame(t(health_pm_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)

>>>>

---
title: 'EldersAIR: data merging for analysis dataset'
author: "Ethan Walker"
date: "Started 11 March 2021, Updated 7 May 2021"
output:
  powerpoint_presentation: default
---

@@

#elders_at_home_log_clean <- read_rds(paste0(file_path, "Output/elders_at_home_log_clean.rds"))

@@

elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds"))

elders_fitbit_visit <- read_rds(paste0(file_path, "Output/elders_fitbit_visit.rds")) 
```

Merge PM, health, stove data 
One observation per sampling visit (up to 4 obs per adult/home)
```{r}
health_pm_visit <- elders_medvisit_clean %>% 
  full_join(elders_pm_at_home_visit,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(demographics,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_extra_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_hx_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(npt_micropem_clean,
            by = c("area", "home_id_num", "adult_id_char", 
                   "home_winter_id", "winter_id", "sampling_visit")) %>% 
  left_join(elders_stove_grades_clean,
            by = c("area", "home", "home_id_num", "treatment")) %>% 
  left_join(elders_stove_usage_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_fitbit_visit,
            by = c("area", "home_winter_id", "adult_id_num", 
                   "winter_id", "sampling_visit")) %>% 
  left_join(elders_kw_report_clean,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_filter_type,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_temp_rh_clean,
            by = c("area", "home_winter_id", "sampling_visit")) %>% 
  left_join(elders_home_activity_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_home_char_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(elders_moisture_clean,
            by = c("area", "home_winter_id", "sampling_visit"))

elders_cleaned_dataset_by_visit <- health_pm_visit %>% 
  mutate(home_floors = as.factor(home_floors),
         home_floors_2level = fct_collapse(home_floors,
                                           "1" = "1",
                                           "2+" = c("2", "3")),
         home_sqm = home_sqft*0.0929,
         home_sqm_2level = cut(home_sqm, breaks = c(0, 112, 10000),
                               labels = c("<111", "111+")),
         home_bedrooms = as.numeric(home_bedrooms),
         home_year_built_2level = cut(home_year_built, 
                                      breaks = c(0, 1981, 10000),
                                      labels = c("<1981", "1981+")),
         home_pets = as.factor(home_pets),
         home_pets_3level = fct_collapse(home_pets,
                                         "0" = "0",
                                         "1" = "1",
                                         "2+" = c("2", "3", "4", "5", "6", "7", "10")),
         income = as.factor(income),
         income_3level = fct_collapse(income,
                                      "< $20,000" = "< $20,000",
                                      "$20,000 to $39,999" = c("$20,000 to $29,999", "$30,000 to $39,999"),
                                      "$40,000+" = c("$40,000 to $49,999", "$50,000 to $74,999", 
                                                     "$75,000 to $99,999", ">= $100,000")),
         stove_age = factor(stove_age,
                            levels = c("0-5 years old", "6-10 years old", "11-15 years old",
                                       "16 + years old", "Do not know")),
         chimney_clean = factor(chimney_clean,
                            levels = c("Less than 6 months ago", "6-12 months ago", 
                                       "12-18 months ago", "More than 18 months ago")),
         burn_level = factor(burn_level,
                            levels = c("No burning", "Light burning", 
                                       "Average burning", "Heavy burning")),
         wood_collect = factor(wood_collect,
                            levels = c("Less than 1 week", "Between 1 week and 1 month", 
                                       "Between 1 month and 3 months", "Between 3 months and 6 months",
                                       "Between 6 months and 1 year", "More than 1 year"))) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) %>% 
  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
  # Filter type isn't really feasible to use in the analysis due to this small sample size
  mutate(treatment_filter_type = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
                                         treatment_filter_type),
         treatment_filter_type = factor(treatment_filter_type2,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix")),
         filter_type = as.factor(filter_type),
         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
  mutate(treatment = as.character(treatment),
         perc_exp_quart = cut(kw_perc_expected,
                                  breaks = c(0, 25, 46, 71, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment == "Filter",
                                     perc_exp_quart, 
                                     treatment),
         treatment = factor(treatment,
                            levels = c("Placebo", "Education", "Filter")),
         filter_compliance = factor(filter_compliance,
                                    levels = c("Placebo", "Education",
                                               "0-25 compliance", "25-50 compliance",
                                               "50-75 compliance", "75-100 compliance"))) %>% 
  select(-treatment_filter_type2) %>% 
  mutate(income_2level = fct_collapse(income,
                                      "20,000 +" = c("$20,000 to $29,999", "$30,000 to $39,999",
                                                     "$40,000 to $49,999", "$50,000 to $74,999",
                                                     "$75,000 to $99,999", ">= $100,000")),
         burn_level_3level = fct_collapse(burn_level,
                                          "No/light burning" = c("No burning", "Light burning")),
         stove_age_2level = fct_collapse(stove_age,
                                         "0-10 years old" = c("0-5 years old", "6-10 years old"),
                                         "11+ years old" = c("11-15 years old", "16 + years old")),
         chimney_clean_3level = fct_collapse(chimney_clean,
                                             "12+ months ago" =
                                               c("12-18 months ago", "More than 18 months ago")),
         wood_collect_3level = fct_collapse(wood_collect,
                                            "< 1 month" = c("Less than 1 week", "Between 1 week and 1 month"),
                                            "1 to 6 months" = c("Between 1 month and 3 months", "Between 3 months and 6 months"),
                                            "6+ months" = c("Between 6 months and 1 year", "More than 1 year")),
         age_2level = cut(age, breaks = c(50, 68, 100),
                          labels = c("50 to 68", "69+")),
         bmi_2level = cut(bmi, breaks = c(0, 29, 100),
                          labels = c("<30", "30+")),
         hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
         steps_4level = cut(steps, breaks = c(0, 1312, 3814, 6734, 31951),
                            labels = c("Q1", "Q2", "Q3", "Q4")),
         steps_2level = cut(steps, breaks = c(0, 3814, 32000),
                            labels = c("<3815", "3815+"))) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(sum_act_smoking = sum(home_act_smoking, na.rm = T),
         sum_act_candle = sum(home_act_candle, na.rm = T),
         sum_act_door = sum(home_act_door, na.rm = T),
         sum_act_gas = sum(home_act_gas, na.rm = T),
         sum_act_elect = sum(home_act_elect, na.rm = T),
         sum_act_propane = sum(home_act_propane, na.rm = T),
         sum_act_oil = sum(home_act_oil, na.rm = T),
         sum_act_incense = sum(home_act_incense, na.rm = T),
         sum_act_kerosene = sum(home_act_kerosene, na.rm = T),
         sum_act_windows = sum(home_act_windows, na.rm = T),
         sum_act_sweep = sum(home_act_sweep, na.rm = T)) %>% 
  group_by(adult_id_char) %>% 
  arrange(adult_id_char, sampling_date) %>% 
  mutate(first_date = first(sampling_date),
         cohort = round_date(first_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  select(-trash) %>% 
  ungroup()

elders_distinct <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = T)

median(elders_distinct$home_year_built, na.rm = T)
table(elders_distinct$sum_act_candle)
```

Save final dataset
```{r}
write_rds(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))

write_csv(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")

# save variable names to use in data dictionary
save_var_names <- data.frame(t(elders_cleaned_dataset_by_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)

Merge PM and at-home data and summarize by sampling visit
## This step is done and data is loaded above
```{r}
elders_at_home_log <- elders_at_home_log_clean %>% 
  select(-home_winter_id, -home, -treatment)

@@
